---
title: "Homework 3"
author: "Christopher Crowe"
date: "October 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

## Problem 1

The following code chunk loads the BRFSS data and conducts the following steps:

* clean variable names
* focus on the "Overall Health" topic
* include only responses from "Excellent" to "Poor"
* organize responses as a factor taking levels from "Excellent" to "Poor"

```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("brfss_smart2010")

brfss_data = brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>%
  mutate(response = ordered(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

Using this cleaned dataset, we can answer/perform the following questions/tasks of interest with the following code chunks: 

### Part A

```{r}
brfss_data %>% 
  filter(year == 2002) %>% 
  distinct(locationabbr, locationdesc) %>% 
  group_by(locationabbr) %>% 
  summarize(observed_locations = n()) %>% 
  filter(observed_locations == 7) %>% 
  knitr::kable()
```

* Q: In 2002, which states were observed at 7 locations?

* A: According to the table above, Connecticut, Florida, and North Carolina each have observations from 7 distinct locations in 2002.


### Part B
  
```{r}
brfss_data %>% 
  filter(year >= 2002 & year <= 2010) %>% 
  group_by(locationabbr, year) %>% 
  summarize(observations = sum(sample_size)) %>% 
  ggplot(aes(x = year, y = observations, color = locationabbr)) +
    geom_line(size = 0.1) +
    labs(
      title = "Observation Plot",
      x = "Year",
      y = "Number of Observations",
      caption = "Data from BRFSS"
    ) +
    viridis::scale_color_viridis(
      name = "State",
      discrete = TRUE
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

* Q: Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

* A: The above "spaghetti plot" shows how many observations there were in each state by year, from 2002 to 2010. However, this graph is not very informative because it is difficult to distinguish the colors of the lines. While we we can that one state in particular had an unusually high number of observations in 2007, it is hard to tell which state corresponds to this color. Additionally, because there are so many lines on top of each other at the bottom of the graph, it is extremely difficult to distinguish one line from another. 

### Part C

```{r}
brfss_data %>% 
  filter(locationabbr == "NY" & response == "Excellent" & (year == 2002 | year == 2006 | year == 2010)) %>% 
  mutate(proportion_excellent = data_value / 100) %>% 
  group_by(year) %>% 
  summarize(mean_proportion_excellent = mean(proportion_excellent), sd_proportion_excellent = sd(proportion_excellent)) %>% 
  knitr::kable()
```

* Q: Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
 
 * A: From this table, we can see that the mean proportion of individuals in NY with excellent overall health has decreased from 0.24 in 2002 but has remained somewhat constant around 0.22 since 2006. Additionally, the standard deviation has steadily decreased by a small amount from 2002 to 2010.

### Part D



```{r}

brfss_data %>% 
  mutate(proportion = data_value / 100) %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(mean_proportion = mean(proportion)) %>% 
  ggplot(aes(x = year, y = mean_proportion, color = locationabbr)) +
    geom_point() +
    scale_x_continuous(breaks = c(2003, 2006, 2009)) +
    facet_grid(~response) +
    labs(
      title = "Trends in Response to Overall Health Question by State 2002 - 2010",
      x = "Year", 
      y = "Mean Proportion",
      caption = "Data from BRFSS"
    ) +
    viridis::scale_color_viridis(
      name = "State",
      discrete = TRUE
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

* Q: For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

* A: In the above graphs, it is still hard to distinguish which data point corresponds to which state, but this graphical representation is successful in showing differences in the overall mean proportions of responses. For example, we can see that across all years and states, there were relatively few "Poor" responses and a much larger proportion of "Very good" responses. We can also see that on average, the proportion of "Poor" responses had less variation (ranging from ~0.0 to ~0.1 for "Poor" vs. ~0.275 to ~4.25 for "Very good"). Additionally, we can identify the presence of outliers even if it may be difficult to discern the state (e.g. it looks like either WY or WV had a much lower proportion of excellent reponses in 2005 compared to all other states).

## Problem 2

```{r, include  = FALSE}
library(p8105.datasets)
data("instacart")

## testing code to be used for inline code below

## count number of rows and columns
nrow(instacart)
ncol(instacart)

## extract example values from cells for key variables
instacart %>% 
  select(user_id) %>% 
  head(1)

instacart %>% 
  select(product_name) %>% 
  head(1)

instacart %>% 
  select(department) %>% 
  head(1)

```

The `instacart` dataset consists of `r nrow(instacart)` rows and `r ncol(instacart)` columns. This dataset contains a variety of information regarding food items sold in a store and relevant order details. The dataset is structured so that each row represents a unique combination of `order_id` and `product_id`. In other words, each row represents one product that was included in a given order, and the contents of the entire order span numerous rows that each begin with the same `order_id`. In addition to these two important variables, other variables of interest may include `user_id` (e.g., `r instacart %>% select(user_id) %>% head(1)`, a unique ID given to the customer who placed the order), `product_name` (e.g., `r instacart %>% select(product_name) %>% head(1)`, the full name of the product), and `department` (e.g., `r instacart %>% select(department) %>% head(1)`, the department of the store where the product can be found).

Using the `instacart` dataset, we can answer/do the following questions/tasks with the following code chunks.

### Part A

```{r, eval = FALSE}
instacart %>% 
  distinct(aisle_id) %>% 
  nrow()

instacart %>% 
  group_by(aisle_id, aisle) %>% 
  summarize(order_count = n()) %>% 
  arrange(desc(order_count)) %>% 
  head(5) %>% knitr::kable()
```

* Q: How many aisles are there, and which aisles are the most items ordered from?
* A: There are `r instacart %>% distinct(aisle_id) %>% nrow()` aisles. The table below shows the aisle ID and name from which the most items are ordered (top 5). `r instacart %>% group_by(aisle_id, aisle) %>% summarize(order_count = n()) %>% arrange(desc(order_count)) %>% head(5) %>% knitr::kable()`

### Part B

```{r}
instacart %>% 
  group_by(aisle_id, department) %>% 
  summarize(order_count = n()) %>% 
  ggplot(aes(x = aisle_id, y = department)) +
    geom_point(aes(size = order_count)) +
    scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140)) +
    labs(
     title = "Order Count by Aisle",
     x = "Aisle ID",
     y = "Department"
    ) +
    theme_bw() +
    theme(legend.position = "bottom")
```

* Q: Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

* A: The above plot attempts to display the number of items ordered in each aisle in an understandable format. On this plot, The `order_count` is indicated by the size of the data point, with larger data points representing a larger number of orders In addition to this, since `aisle_id` provides little information on its own, the corresponding `department` is listed on the y-axis. Using this layout, we can, for example, see that most orders are coming from Aisles 24 and 83 in the produce department. We can also see which aisle, within a given department, has the largest number of orders. 

### Part C

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(product_name, aisle) %>% 
  summarize(order_count = n()) %>% 
  arrange(desc(order_count)) %>% 
  group_by(aisle) %>% 
  top_n(n = 1, wt = order_count) %>% 
  rename(top_product_name = product_name) %>% 
  select(aisle, top_product_name, order_count) %>% 
  arrange(aisle) %>% 
  knitr::kable()

```

* Q: Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

* A: According to the table above, the most popular items in the "baking ingredients", "dog food care", and "packaged vegetables fruits" aisles are Light Brown Sugar, Snack Sticks Chicken & Rice Recipe Dog Treats, and Organic Baby Spinach, respectively. 

### Part D

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_order_time = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_order_time) %>% 
  knitr::kable()
  
```

Q: Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

A: The above table is useful in comparing the times at which Pink Lady Apples and Coffee Ice Cream are ordered. For example, we can see that Pink Lady Apples are typically ordered earlier than Coffee Ice Cream. However, the table could be improved if we knew how days of the week were coded. As it is, we do not know which number corresponds to which day of the week, though we may assume Day 0 is either Sunday or Monday. 

## Problem 3

```{r, include = FALSE}
library(p8105.datasets)
data("ny_noaa")

## testing code to be used for inline code below

## count number of rows and columns
nrow(ny_noaa)
ncol(ny_noaa)

## extract example values from cells for key variables
ny_noaa %>% 
  select(id) %>%
  head(1)

ny_noaa %>% 
  select(date) %>% 
  head(1)

ny_noaa %>% 
  select(prcp) %>% 
  filter(prcp != "NA") %>%
  head(1)

ny_noaa %>% 
  select(tmax) %>% 
  filter(tmax != "NA") %>%
  head(1)

## count the number of rows where something is missing
ny_noaa %>% 
  filter(is.na(tmax) | is.na(tmin) | is.na(prcp) | is.na(snow) | is.na(snwd)) %>% 
  nrow()

```

The `ny_noaa` dataset consists of `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. This dataset contains a variety of weather information recorded by five weather stations across New York state. The dataset is structured so that each row represents the weather information for a particular weather station on any given day. Examples of key variables include  `id` (e.g., `r ny_noaa %>% select(id) %>% head(1)`, a unique ID given to each of the five NOAA weather stations in New York state), `date` (e.g., `r ny_noaa %>% select(date) %>% head(1)`, the date on which the data was recorded), `prcp` (e.g., `r ny_noaa %>% select(prcp) %>% filter(prcp != "NA") %>% head(1)`, the amount of precipitation in tenths of a millimeter), and `tmax` (e.g., `r ny_noaa %>% select(tmax) %>% filter(tmax != "NA") %>% head(1)`, the maximum temperature in tenths of a degree Celsius). This dataset has a large number of missing values, with `r ny_noaa %>% filter(is.na(tmax) | is.na(tmin) | is.na(prcp) | is.na(snow) | is.na(snwd)) %>% nrow()` observations where at least one piece of weather information is missing. 

Using the `ny_noaa` dataset, we can answer/do the following questions/tasks with the following code chunks.